Winter weather on exam dates and matriculation for a prestigious university in Japan

### 2021-04-13 Writeups/paper.tex
Revision before sending to Laura

\maketitle
\begin{abstract}
  \singlespacing
     \noindent

@@

The econometric analyses show that low temperature and snow cumulated on the ground on exam dates reduce the matriculation share.
The estimates are not only statistically but also economically significant.
Changing the temperature on the exam dates from 3-6 $^o$C to lower than 0 $^o$C decreases the matriculation share by 0.18\%, which amounts to 9\% of the average or 22\% of the median matriculation shares.
The results provide implications for effective measures to be taken to mitigate the negative impact of winter weather.
These include preparing waiting rooms for exam takers to warm up for equalizing their pre-exam conditions or adjusting the importance of the scores at Center Test in the admission process.

This study mainly contributes to the recent but growing literature of the effects of climate factors on the outcome of high-stakes exams.
Several previous studies have found the adverse effect of high temperature on learning, which reduces performance at exams \citep{Park2020b, Cho2017}.
More closely to my study, \citet{GraffZivin2020} and \citet{Park2020a} show that heat on exam dates decreases the exam scores at high-stakes exams in China and the U.S., respectively.
Considering the importance of the exam results for later life outcomes, it is important to understand how external factors such as temperature affect exam performance and what can be done to ameliorate such effects.
To complement the literature, I investigate the effect of \textit{low} temperature on exam performance.

@@

This study proceeds as follows.
In Section \ref{sec:background}, the context of this study is introduced.
Section \ref{sec:data} describes the data. 
I introduced my empirical strategy in Section \ref{sec:empirical_strategy}.
In Section \ref{sec:results}, I present the empirical results, and Section \ref{sec:conclusion} concludes.

@@

In Japan, for admission to public universities, students are usually required to take two exams.
The first-stage test is called Center Test.
Students take an exam on consecutive two days (Saturday and Sunday) in mid-January.
Most public universities, including UTokyo, require students to take exams on social science, language, and English on the first day and math and natural science on the second day.
The exams start at 9:30 AM on both days and end at 6:10 PM on the first day and 5:50 PM on the second day.
Questions are multiple-choice and common for all exam takers across Japan.
High-school seniors take the test at designated sites in the exam area\footnote{ \label{footnote:prefecture}
  Typically, an exam area coincides with a prefecture (i.e. there is one exam area in one prefecture).
  Large prefectures such as Hokkaido and Okinawa have several exam areas.
  Also, several municipalities in a prefecture can be sometimes considered to be a part of a neighboring prefecture.
  For instance, students in a couple of municipalities in Saitama prefecture could take Center Test in Tokyo.
} where their high schools are located.
On the other hand, high-school graduates are assigned to test sites in the exam area where the homes used in their applications are located.\footnote{
  \citet{Watanabe2013} provides detailed descriptions on the administrations of Center Test.
}

Since January is the middle of winter in Japan, heavy snow can affect public transportation on exam dates.
This sometimes makes it difficult for exam takers to arrive at exam sites on time.
For the affected students, the following accommodations can be provided:
(i) delaying the start of exams by a few hours for all exam takers at a test site,
(ii) allowing the affected students to take exams in a separate room, and
(iii) allowing the affected students to take make-up exams a week later.
The same accommodations can be given to students who could not take exams for other reasons such as sickness or accidents.

@@

In admission for UTokyo, scores at Center Test are used for two purposes.
First, the scores are used for eligibility to take the second-stage exam.
Thresholds varying across years and across majors are set, and if scores at Center Test are below the thresholds, students cannot take the second-stage exam.
%Since the thresholds are announced after the application, exam takers do not know if their scores are above or below the thresholds.
Secondly, the scores at Center Test are taken into consideration for admission.
The final score of a student is calculated by the weighted sum of exam scores at the first- and second-stage exams.\footnote{
  In 2020, for the first-period exam takers, the weights for Center Test and the second-stage exam are 20\% and 80\%.
  %Also, until 2015, for the third-period exam takers, scores at Center Test were used only for eligibility and were not accounted for the final admission.
}

@@

The summary statistics are provided in Table \ref{tab:sum_stat}.
While the mean of the matriculation share (\%) is $2.08$,\footnote{
  Since there are matriculations from abroad, the mean value does not coincide with $1 / 47 \approx 2.12\%$.
  I do not consider matriculations from abroad in the analyses since their admission is mostly based on interviews and essays.
} the distribution is starkly skewed.
As shown in Figure \ref{fig:admission_map}, many prefectures have less than 1\% shares.
Also, the matriculation shares of the top 5 prefectures amount to more than 60\%.
This results in a small median matriculation share, $0.79 \%$.

@@

%\begin{figure}[H]
%  \centering
%  \caption{Map of matriculation shares (\%) (average across years)}
%  \includegraphics[width = 0.7\textwidth]{../Output/images/admission_map.png}
%  \label{fig:admission_map}
%  \footnotesize
%  \begin{tablenotes}
%    \item Notes:
%      This map shows the matriculation shares (\%) for the University of Tokyo from each prefecture.
%      The matriculation share is calculated based on the ratio of the number of students from high schools in the prefecture to the total number of matriculation.
%      Average of the shares across years is shown in the map.
%      Although not shown in the legend due to the lack of the space, the lightest color in the map means the matriculation share between 0 to 0.5\% 
%  \end{tablenotes}
%\end{figure}

@@

The objective of the analysis is to analyze the effect of temperature and other weather variables on exam dates on the matriculation share for UTokyo.
The reason why I use the matriculation share for UTokyo is as follows:
consider a matriculation share for a middle-ranked university.
Suppose also that low temperature has a negative impact on exam scores.
Then, 
(i) students from the prefecture who would get into the middle-ranked university if the temperature was higher could go to a lower-ranked university, but
(ii) students from the prefecture who would get into a higher-ranked university may get into the middle-ranked university.
Hence, the matriculation share of the prefecture for the middle-ranked university may not be affected by low temperature.
However, the second effect is unlikely to be large for UTokyo since it is known to be the best university in Japan.\footnote{
  \citet{Kawaguchi2008} describes that ``The University of Tokyo has occupied the top position of the single-peaked university hierarchy since its establishment in 1877.''
}
%Secondly, since UTokyo is a national university, the tuition is cheaper than most private universities.\footnote{
%  According to \citet{mext}, in 2018, the annual tuition for national universities was 538,633 JPY ($\approx$ 4,900 USD), while the average annual tuition for private university was 904,146 ($\approx$ 8,200 USD).
%  %https://www.mext.go.jp/a_menu/koutou/shinkou/07021403/1412031_00001.htm
%}
Therefore, I rule out the second possibility and attribute a reduction in matriculation shares to the worse exam performance.

@@

One potential problem in the analysis is that a student from a high school in a prefecture does not necessarily take Center Test in the same municipality.
This means that, for these students, the weather at the exam sites is measured with errors.
The primary reason for this to happen is that a senior student at a high school located at a prefecture different from his home prefecture fails to get into any university for the first year and retake Center Test next year in his home prefecture.
%\footnote{
%  In 2020, less than 40\% of the admitted students are high-school graduates.
%  In the same year, according to the School Basic Survey, the share of high-school students from junior high schools in different prefectures was 3.9\% \citep{eStat}.
%  If I assume that that this number corresponds to the share of high-school students whose hometowns are in different prefectures, I obtain a rough estimate of the share of the high-school graduates whose high schools and hometowns are in different prefectures as $0.40 * 0.039 = 1.6\%$. 
%}
I consider that, conditional on prefecture fixed effects, the measurement error is likely to be uncorrelated with the error term.
Hence, this classical measurement error can make the estimates conservative.
Another potential concern is that, as discussed in footnote \ref{footnote:prefecture}, students in several municipalities in a prefecture could take Center Test in a neighboring prefecture.
Since neighboring prefectures are geographically close, I consider that this measurement error is not large.

@@

The main empirical results are provided in Table \ref{tab:main_reg}.
The standard errors are clustered at the prefecture level.
Column (1) shows the result with a linear specification.
While the point estimate is positive, which means matriculation shares are higher if an exam is held on warmer days, it is statistically insignificant. 
The estimates with a non-linear specification allowing for flexible temperature effects (column (2)) show a negative and statistically significant impact of cold weather. 
Since the mean matriculation share is $2.08$\%, by reducing the temperature from the reference bin (between 3 and 6 $^o$C) to the lowest temperature bin, the matriculation share decreases by 9\%.
Moreover, due to the stark skewness in the distribution of matriculation shares, the median matriculation share is $0.79$\%.
Therefore, the same temperature change decreases the matriculation share by 22\% of the median value.
Also, the plateau effect of temperature above 6 $^o$C could be the reason why the linear specification in column (1) could not capture the non-linear effect of temperature and why the coefficient is statistically insignificant.
Although for a quite different temperature range, similar plateau effects are found in \citet{Park2020a}.

In columns (3) and (4), I add rainfall and snowfall in the regressions.
While temperature coefficients are barely affected, I find that rainfall and snowfall on the exam dates do not impact the matriculation shares.
In columns (5) and (6), instead of snowfall on the exam dates, I use cumulated snow on the ground and find its negative impact on the outcome. 
To capture the effect of the existence of a certain amount of snow on the ground, column (7) allows a non-linear effect of cumulated snow.
The result shows that snow on the ground cumulated more than 10 cm ($\approx$ 4 inches) decreases the matriculation share by 0.11 percentage point.\footnote{
  Following the falsification test in \citet{Cho2017}, I check if weather variables one year after the actual Center Test affect the matriculation shares.
  The results in Table \ref{tab:reg_placebo_exam} show not only that the weather one year after Center Test has insignificant impacts on matriculation shares but also including the weather variables one year after Center Test barely change the coefficients on contemporaneous weather variables.
}

@@

In Table \ref{tab:reg_pre10}, the coefficients on snowfall and cumulated snow on the ground are statistically significant.
Combined with the finding in Table \ref{tab:main_reg} that cumulated snow on the ground on the exam dates decreases students' performance, two potential explanations can be given:
(i) snow before the exams can adversely affect students' preparation, and
(ii) snowfall before the exams cumulates snow on the ground and students' performance is affected by the cumulated snow on the exam dates.
Due to the high correlation of cumulated snow on the exam dates and 10 days before the exams, I cannot identify which mechanism is dominant.

@@

The results have implications for the admission process of universities.
Recognizing the limitation to measure students' cognitive ability through an exam affected by external factors such as weather, universities may use other measures to select students.
Examples include extracurricular activities and personal statements, which are actively utilized in other countries such as the United States.
Also, they may change the weights for Center Test and the second-stage exam.
Since the second-stage exams are held at each university, all exam takers are exposed to similar environments.
This rules out the possibility that exam takers at different places are affected differently by their conditions.
Given the importance of exam results for later life outcomes, it is desirable that universities take measures to mitigate the effect of nuisance factors on admission.

@@

\begin{figure}[H]
  \centering
  \caption{Map of matriculation shares (\%) (average across years)}
  \includegraphics[width = 0.7\textwidth]{../Output/images/admission_map.png}
  \label{fig:admission_map}
  \footnotesize
  \begin{tablenotes}
    \item Notes:
      This map shows the matriculation shares (\%) for the University of Tokyo from each prefecture.
      The matriculation share is calculated based on the ratio of the number of students from high schools in the prefecture to the total number of matriculation.
      Average of the shares across years is shown in the map.
      Although not shown in the legend due to the lack of the space, the lightest color in the map means the matriculation share between 0 to 0.5\% 
  \end{tablenotes}
\end{figure}

>>>>

\maketitle
\begin{abstract}
  \singlespacing
    \noindent 
    Whereas there are growing numbers of studies on how heat affects cognitive performance at high-stakes settings, the effect of cold weather has been less studied.
    To fill this gap, I analyze the effects of winter weather on exam dates on performance at the centralized exam for university admission in Japan.
    I find that low temperature and snow cumulated on the ground reduce matriculation shares for the most prestigious university in Japan, the University of Tokyo.
    This result suggests that the winter weather on exam dates worsens exam performance. 
    The implications on how to design university admissions are discussed.

@@

The econometric analyses show that low temperature and snow cumulated on the ground on exam dates reduce the matriculation share.
The estimates are not only statistically but also economically significant.
Changing the temperature on the exam dates from 3-6 $^o$C to below 0 $^o$C decreases the matriculation share by 0.18\%, which amounts to 9\% of the average or 22\% of the median matriculation shares.
The results provide implications for effective measures to be taken to mitigate the negative impact of winter weather.
These include preparing waiting rooms for exam takers to warm up for equalizing their pre-exam conditions or adjusting the importance of the scores at Center Test in the admission process.

This study mainly contributes to the recent but growing literature of the effects of climate factors on the outcome of high-stakes exams.
Several previous studies have found the adverse effect of high temperature on learning, which exacerbates performance at exams \citep{Park2020b, Cho2017}.
More closely to my study, \citet{GraffZivin2020} and \citet{Park2020a} show that heat on exam dates decreases the scores at high-stakes exams in China and the U.S., respectively.
Considering the importance of the exam results for later life outcomes, it is important to understand how external factors such as temperature affect exam performance and what can be done to ameliorate such effects.
To complement the literature, I investigate the effect of \textit{low} temperature on exam performance.

@@

This study proceeds as follows.
In Section \ref{sec:background}, the context of this study is introduced.
Section \ref{sec:data} describes the data. 
I introduce my empirical strategy in Section \ref{sec:empirical_strategy}.
In Section \ref{sec:results}, I present the empirical results, and Section \ref{sec:conclusion} concludes.

@@

In Japan, for admission to public universities, students are usually required to take two exams.
The first-stage test is called Center Test.
Students take an exam on consecutive two days (Saturday and Sunday) in mid-January.
Most public universities, including UTokyo, require students to take exams on social science, language, and English on the first day and mathematics and natural science on the second day.
The exams start at 9:30 AM on both days and end at 6:10 PM on the first day and 5:50 PM on the second day.
Questions are multiple-choice and common for all exam takers across Japan.
High-school seniors take the test at designated sites in the exam area\footnote{ \label{footnote:prefecture}
  Typically, an exam area coincides with a prefecture (i.e. there is one exam area in one prefecture).
  Large prefectures such as Hokkaido and Okinawa have several exam areas.
  Also, several municipalities in a prefecture can be sometimes considered to be a part of a neighboring prefecture.
  For instance, students in a couple of municipalities in Saitama prefecture could take Center Test in Tokyo.
} where their high schools are located.
On the other hand, high-school graduates are assigned to test sites in the exam area where the homes used in their applications are located.\footnote{
  \citet{Watanabe2013} provides detailed descriptions on the administrations of Center Test.
}

Since January is the middle of winter in Japan, heavy snow can affect public transportation on exam dates.
This sometimes makes it difficult for exam takers to arrive at exam sites on time.
For the affected students, the following accommodations can be provided:
(i) delaying the start of exams by a few hours for all exam takers at a test site,
(ii) allowing the affected students to take exams in a separate room, and
(iii) allowing the affected students to take make-up exams a week later.
These accommodations can be given to students who could not take exams for other reasons such as sickness or accidents.

@@

In admission for UTokyo, scores at Center Test are used for two purposes.
First, the scores are used for eligibility to take the second-stage exam.
Thresholds varying across years and majors are set, and if scores at Center Test are below the thresholds, the students cannot take the second-stage exam.
%Since the thresholds are announced after the application, exam takers do not know if their scores are above or below the thresholds.
Secondly, the scores at Center Test are taken into consideration for admission.
The final score of a student is calculated by the weighted sum of exam scores at the first- and second-stage exams.\footnote{
  In 2020, the weights for Center Test and the second-stage exam are 20\% and 80\%.
  Faculty of Medicine additionally conducts interviews on the next day of the second-stage exam to screen applicants based on community skills and personal maturity.
  %Also, until 2015, for the third-period exam takers, scores at Center Test were used only for eligibility and were not accounted for the final admission.
}

@@

The summary statistics are provided in Table \ref{tab:sum_stat}.
Whereas the mean matriculation share is $2.08 \%$, the median share is much smaller ($0.79 \%$).\footnote{
  Since there are matriculations from foreign high schools, the mean value does not coincide with $1 / 47 \approx 2.12\%$.
  %I do not consider matriculations from foreign high schools in the analyses since I do not know where the students take Center Test.
  %Also, the majority of the students from foreign high schools are admitted mostly based on interviews and essays.
  I do not consider matriculations of students from foreign high schools in the analyses since the majority of the students are admitted through a process called Special Screening Tests, which does not require Center Test scores.
}
This is because the distribution of the matriculation share is starkly skewed:
while the matriculation shares of the top 5 prefectures amount to more than 60\%, many prefectures have less than 1\% shares (Figure \ref{fig:admission_map}).

@@

%\begin{figure}[H]
%  \centering
%  \caption{Map of matriculation shares (\%)}
%  \includegraphics[width = 0.7\textwidth]{../Output/images/admission_map.png}
%  \label{fig:admission_map}
%  \footnotesize
%  \begin{tablenotes}
%    \item Notes:
%      This map shows the matriculation shares (\%) for the University of Tokyo from each prefecture.
%      The matriculation share is calculated based on the ratio of the number of students from high schools in the prefecture to the total number of matriculation.
%      Average of the shares across years is shown in the map.
%      Although not shown in the legend due to the lack of the space, the lightest color in the map means the matriculation share between 0 to 0.5\% 
%  \end{tablenotes}
%\end{figure}

@@

I analyze the effect of temperature and other weather variables on exam dates on the matriculation share for UTokyo.
The reason why I use the matriculation share for UTokyo is as follows:
consider a matriculation share for a middle-ranked university.
Suppose also that low temperature has a negative impact on exam scores.
Then, 
(i) students in the prefecture who would get into the middle-ranked university if the temperature was higher could go to a lower-ranked university, but
(ii) students in the prefecture who would get into a higher-ranked university may get into the middle-ranked university.
Hence, the matriculation share of the prefecture for the middle-ranked university may not be affected by low temperature.
However, the second effect is unlikely to be large for UTokyo since it is known to be the best university in Japan.\footnote{
  \citet{Kawaguchi2008} describes that ``The University of Tokyo has occupied the top position of the single-peaked university hierarchy since its establishment in 1877.''
}
%Secondly, since UTokyo is a national university, the tuition is cheaper than most private universities.\footnote{
%  According to \citet{mext}, in 2018, the annual tuition for national universities was 538,633 JPY ($\approx$ 4,900 USD), while the average annual tuition for private university was 904,146 ($\approx$ 8,200 USD).
%  %https://www.mext.go.jp/a_menu/koutou/shinkou/07021403/1412031_00001.htm
%}
Therefore, I rule out the second possibility and attribute a reduction in matriculation shares to the worse performance at Center Test.

@@

One potential problem in the analysis is that a student from a high school in a prefecture does not necessarily take Center Test in the same municipality.
This means that, for these students, the weather at the exam sites is measured with errors.
The primary reason for this to happen is that a senior student at a high school located at a prefecture different from his home prefecture fails to get into any university and retake Center Test next year in his home prefecture.
In 2020, 38\% of the admitted students are high-school graduates.
Moreover, in the same year, according to the School Basic Survey, the share of high-school students from junior high schools in different prefectures was 3.9\% \citep{eStat}.
%Hence, if I assume that that this number corresponds to the share of high-school students whose hometowns are in different prefectures, I obtain a rough estimate of the share of the high-school graduates whose high schools and hometowns are in different prefectures as $0.40 * 0.039 = 1.6\%$. 
Hence, if I use this number as an approximation for the share of high-school students whose hometowns are in different prefectures, I obtain a rough estimate of the share of students whose exam-date weather is measured with errors for the reason above as $0.38 * 0.039 = 1.5\%$. 
Given this small proportion of exam takers who are susceptible to the measurement error, I consider that the bias on the estimates is negligible.
%I consider that, conditional on prefecture fixed effects, the measurement error (i.e. the difference in temperatures on the exam dates between the exam prefecture and home prefecture) is likely to be uncorrelated with the error term.
%Hence, this classical measurement error can make the estimates conservative.
Another potential concern is that, as discussed in footnote \ref{footnote:prefecture}, students in several municipalities in a prefecture could take Center Test in a neighboring prefecture.
Since neighboring prefectures are geographically close, I consider that the magnitude of the measurement error is small.

@@

The main empirical results are provided in Table \ref{tab:main_reg}.
The standard errors are clustered at the prefecture level.
Column (1) shows the result with a linear specification.
While the point estimate is positive, which means matriculation shares are higher if an exam is held on warmer days, it is statistically insignificant. 
The estimates with a non-linear specification allowing for flexible temperature effects (column (2)) show a negative and statistically significant impact of cold weather. 
Since the mean matriculation share is $2.08$\%, by reducing the temperature from the reference bin (between 3 and 6 $^o$C) to the lowest temperature bin (below 0 $^o$C), the matriculation share decreases by 9\%.
Moreover, due to the stark skewness in the distribution of matriculation shares, the median matriculation share is $0.79$\%.
Therefore, the same temperature change decreases the matriculation share by 22\% of the median value.
Also, the plateau effect of temperature above 6 $^o$C could be the reason why the linear specification in column (1) could not capture the non-linear effect of temperature and the estimate is statistically insignificant.
Although for a quite different temperature range, similar plateau effects are found in \citet{Park2020a}.

In columns (3) and (4), I add rainfall and snowfall in the regressions.
While temperature coefficients are barely affected, I find that rainfall and snowfall on the exam dates do not impact the matriculation shares.
In columns (5) and (6), instead of snowfall on the exam dates, I use cumulated snow on the ground and find its negative impact on the outcome. 
To capture the effect of the existence of a certain amount of snow on the ground, column (7) allows a non-linear effect of cumulated snow.
The result shows that snow on the ground cumulated more than 10 cm ($\approx$ 4 inches) decreases the matriculation share by 0.11 percentage point.\footnote{
  Following the falsification test in \citet{Cho2017}, I check if weather variables one year after the actual Center Test affect the matriculation shares.
  The results in Table \ref{tab:reg_placebo_exam} show not only that the weather one year after Center Test has insignificant impacts on matriculation shares but also including the weather variables one year after Center Test barely changes the coefficients on contemporaneous weather variables.
}

@@

In Table \ref{tab:reg_pre10}, the coefficients on snowfall and cumulated snow on the ground are statistically significant.
Combined with the finding in Table \ref{tab:main_reg} that cumulated snow on the ground on the exam dates decreases students' performance, two potential explanations can be given:
(i) snow before the exams can adversely affect students' preparation, and
(ii) snowfall before the exams cumulates snow on the ground and students' performance is negatively affected by the cumulated snow on the exam dates.
Due to the high correlation of cumulated snow on the exam dates and 10 days before the exams, I cannot identify which mechanism is dominant.

@@

The results in this study have implications for the admission process of universities.
Recognizing the limitation to measure students' cognitive ability through an exam affected by external factors such as weather, universities may use other measures to select students.
Examples include extracurricular activities and personal statements, which are actively utilized in other countries such as the United States.
Also, universities may adjust the weights for Center Test and the second-stage exam.
Since the second-stage exams are held at each university, all exam takers are exposed to similar environments.
This rules out the possibility that exam takers at different places are affected differently by their conditions.
Given the importance of exam results for later life outcomes of students, it is desirable that universities take measures to mitigate the effect of seemingly irrelevant factors on admission.

@@

\begin{figure}[H]
  \centering
  \caption{Map of matriculation shares (\%)}
  \includegraphics[width = 0.7\textwidth]{../Output/images/admission_map.png}
  \label{fig:admission_map}
  \footnotesize
  \begin{tablenotes}
    \item Notes:
      This map shows the matriculation shares (\%) for the University of Tokyo from each prefecture.
      The matriculation share is calculated based on the ratio of the number of students from high schools in the prefecture to the total number of matriculation.
      Average of the shares across years is shown in the map.
      Although not shown in the legend due to the lack of the space, the lightest color in the map means the matriculation share between 0 to 0.5\% 
  \end{tablenotes}
\end{figure}
